---
title: "p8105_hw2_as5457"
author: "Sunny Siddique"
date: "October 1, 2018"
output: html_document
---

Problem 1

```{r}
library(tidyverse)
```


#Problem 1


##Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names(dat = .) %>% 
  rename (route_1 = route1, route_2 = route2, route_3 = route3, route_4 = route4, route_5 = route5, route_6 = route6, route_7 = route7, route_8 = route8, route_9 = route9, route_10 = route10, route_11 = route11) %>% 
  gather(key = route, value = route_name, route_1:route_11) %>% 
  separate (route, into = c("route_str", "route_number"), sep = "_") %>% 
    select (line, station_name, station_latitude, station_longitude, route_number, route_name, entry, vending, ada) %>% 
  mutate(entry=ifelse(entry=="YES", yes=TRUE, no=FALSE))
```

##Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

The variables in this dataset provide information regarding the line, name, exact location (latitude/longitude), the name and number of routes available at the station, entry, vending and ADA compliance status. To clean this dataset, I first used the janitor package to change the variable names to lower case and separated by _. Then I renamed the route variables and separated the character variables by underscores to be able to later split it into two variables: route number and route name. I then selected only the variables specified in the prompt and used the mutate and ifthen function to change the character variable "entry" into a logical variable. 


##How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.
The number of distinct stations is `r nrow(distinct(transit_data, station_name, line))`.


##How many stations are ADA compliant?
The number of stations that are ADA compliant is `r mean(transit_data$ada)`.

```{r}
transit_data %>% 
  group_by (station_name, line) %>% 
  tally(wt = ada)

```


What proportion of station entrances / exits without vending allow entrance?
Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?






















