---
title: "p8105_hw2_as5457"
author: "Sunny Siddique"
date: "October 1, 2018"
output: html_document
---

Problem 1

```{r}
library(tidyverse)
```


#Problem 1


##Read and clean the data; retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names(dat = .) %>% 
  rename (route_1 = route1, route_2 = route2, route_3 = route3, route_4 = route4, route_5 = route5, route_6 = route6, route_7 = route7, route_8 = route8, route_9 = route9, route_10 = route10, route_11 = route11) %>% 
  gather(key = route, value = route_name, route_1:route_11) %>% 
  separate (route, into = c("route_str", "route_number"), sep = "_") %>% 
    select (line, station_name, station_latitude, station_longitude, route_number, route_name, entry, vending, ada) %>% 
  mutate(entry=ifelse(entry=="YES", yes=TRUE, no=FALSE))
```

##Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

The variables in this dataset provide information regarding the line, name, exact location (latitude/longitude), the name and number of routes available at the station, entry, vending and ADA compliance status. To clean this dataset, I first used the janitor package to change the variable names to lower case and separated by _. Then I renamed the route variables and separated the character variables by underscores to be able to later split it into two variables: route number and route name. I then selected only the variables specified in the prompt and used the mutate and ifthen function to change the character variable "entry" into a logical variable. 


##How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.
The number of distinct stations is `r nrow(distinct(transit_data, station_name, line))`.


##How many stations are ADA compliant?
The number of stations that are ADA compliant is.

```{r}
transit_data %>% 
  filter (ada == "TRUE") %>% 
  distinct (station_name)

```


##What proportion of station entrances / exits without vending allow entrance?

```{r}
vending =
transit_data %>% 
  filter (vending == "NO" & entry == "TRUE") %>% 
  distinct (station_name)
nrow(vending)/nrow(distinct(transit_data, station_name, line))
```

##Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}
a_train = transit_data %>% 
  filter (route_name == "A") %>% 
  distinct(station_name)
nrow(a_train)
```

```{r}
ada_comp = transit_data %>% 
  filter (route_name == "A", ada == "TRUE") %>% 
  distinct(station_name)
nrow(ada_comp)
```

#Problem 2

Read and clean the Mr. Trash Wheel sheet:

- specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function)
- use reasonable variable names
- omit rows that do not include dumpster-specific data, rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

```{r}
library(readxl)
trash_wheel = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 1, range = "A2:N258") %>% 
  janitor::clean_names(dat = .) %>% 
  filter(!is.na(dumpster))
```

```{r}
mutate(trash_wheel, round(sports_balls, digits = 0)
)
```

Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. 

```{r}
prec17 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 3, range = "A2:B14") %>% 
  janitor::clean_names(dat = .) %>% 
  filter(!is.na(total))
prec17$year = 2017
```

```{r}
prec16 = readxl::read_excel("./data/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = 4, range = "A2:B14") %>% 
  janitor::clean_names(dat = .) %>% 
  filter(!is.na(total))
prec16$year = 2016
```


##Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

```{r}
combined_data = (weather_df, trash_wheel, by = "date")
combined_data
```


Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2017? What was the median number of sports balls in a dumpster in 2016?


















